Analysis-ready VCF at Biobank scale using Zarr

### 2024-05-24 paper.tex
Add note on Genozip
Add QC citations from Eric

Processing of Biobank scale datasets can be split into a 
few broad categories. The most basic analysis 
is quality control (QC). Variant QC is an 
involved and multi-faceted 
task [CITATIONS?], often requiring interactive, exploratory analysis
and incurring substantial computation over multiple QC fields.
Genotype calls are often refined via statistical methods,
% methods~\citep[e.g.][MORE and CHECK]{browning2021fast,browning2023statistical}
for example by phasing~[CITES] and imputation[CITES], 
creating additional dataset copies.
A common task to perform after QC is 
a genome wide association study (GWAS)~\cite{uffelmann2021genome}. 
The majority of tools for 
performing GWAS and related analyses require
data to be in PLINK format~\cite[e.g][]{chang2015second} [MORE],
and so data must be ``hard-called'' according to some QC criteria
and exported to additional copies.
Finally, variation datasets are often queried in exploratory
analyses, to find regions or samples of interest for a particular 
study~\cite[e.g.][]{chen2024novo}.

@@

>>>>

Processing of Biobank scale datasets can be split into a 
few broad categories. The most basic analysis 
is quality control (QC). Variant QC is an 
involved and multi-faceted 
task~\cite{marees2018tutorial,panoutsopoulou2018quality}, 
often requiring interactive, exploratory analysis
and incurring substantial computation over multiple QC fields.
Genotype calls are often refined via statistical methods,
% methods~\citep[e.g.][MORE and CHECK]{browning2021fast,browning2023statistical}
for example by phasing~[CITES] and imputation[CITES], 
creating additional dataset copies.
A common task to perform after QC is 
a genome wide association study (GWAS)~\cite{uffelmann2021genome}. 
The majority of tools for 
performing GWAS and related analyses require
data to be in PLINK format~\cite[e.g][]{chang2015second} [MORE],
and so data must be ``hard-called'' according to some QC criteria
and exported to additional copies.
Finally, variation datasets are often queried in exploratory
analyses, to find regions or samples of interest for a particular 
study~\cite[e.g.][]{chen2024novo}.

@@

Genozip was used under the terms of the free Genozip Academic license.
Genozip was only used on simulated data, in compliance with
the ``No Commercial Data'' criterion.
